Search Strategies for Functional Logic Programming

### 2012-01-17 papers/atps2012/paper.tex

\begin{abstract}
In this paper we discuss our practical experiences
with the use of different search strategies
in functional logic programs.
In particular, we show that complete strategies,
like breadth-first search or iterative deepening search,
are a reasonable alternative to incomplete strategies, like depth-first
search, that have been favored in the past for logic programming languages.
\end{abstract}

@@

Figure~\ref{fig:bench-all-solutions} shows the benchmark results
when all values of a given expression are computed.
All benchmark programs are non-deterministic programs.
For instance, ``PermSort'' sorts
a list containing 15 elements by enumerating all permutations
and selecting the sorted ones,
``Last'' compute the last element \code{x}
of a list \code{xs} containing 100,000 elements
by solving the equation \ccode{ys++[x] =:= xs}
(see Section~\ref{sec:curry}),
``Half'' computes the half \code{y} of a natural number \code{x}
(in Peano representation) by solving the equation \code{y+y=:=x},
\ldots

@@

Complete strategies could be the default!

If efficiency important and DFS safe (e.g., finite search space): use DFS

Future work: approximate safeness of DFS

>>>>

\begin{abstract}
In this paper we discuss our practical experiences
with the use of different search strategies
in functional logic programs.
In particular, we show that complete strategies,
like breadth-first search or iterative deepening search,
are a viable alternative to incomplete strategies, like depth-first
search, that have been favored in the past for logic programming languages.
\end{abstract}

@@

Figure~\ref{fig:bench-all-solutions} shows the benchmark results
when all values of a given expression are computed.
All benchmark programs are non-deterministic programs.
``PermSort'' sorts a list containing 15 elements by enumerating
all permutations and selecting the sorted ones,
``Last'' compute the last element \code{x}
of a list \code{xs} containing 100,000 elements
by solving the equation \ccode{ys++[x] =:= xs}
(see Section~\ref{sec:curry}),
``Half'' computes the half \code{y} of a natural number \code{x}
(in Peano representation) by solving the equation \code{y+y=:=x},
``Embed'' searches for an embedding of a message
which is ``hidden'' in another message
(see \cite[Sect.~4.1]{Antoy10JSC}),
\ldots

@@

We discussed the use of different search strategies
in functional logic programs.
In order to enable user-programmable search strategies,
modern functional logic languages like Curry
provide primitives to represent non-deterministic computations
or values as data structures that can be traversed like
any term structure.
The Curry implementation KiCS2, considered in this paper,
provides a primitive \code{getSearchTree}
that returns a tree representation of a non-deterministic computation.
This representation can be used to define various search strategies,
like depth-first search, breadth-first search, or iterative deepening
search.

We have practically evaluated and compared
the efficiency of these strategies.
The benchmarks indicated that complete strategies
are a viable alternative to incomplete strategies,
that have been favored in the past due to limited memory requirements.
Thus, it could be reasonable to make
complete strategies the default in functional logic languages.
This would have a good impact on teaching declarative programming
to beginners.
Moreover, decoupling non-deterministic programs
from their search strategy could lead to a more declarative
programming style (instead of the use of predicates with
side effects as in Prolog) and enable more potential for
optimization, e.g., parallel search strategies.
Furthermore, if efficiency and memory limitations are important,
one can still use an efficient strategy like depth-first search
provided that it is able to find all solutions
(e.g., in case of a finite search space).
It is an interesting topic for future work
to statically approximate situations where
the use of incomplete strategies is sufficient.
